How Residents of Canada Feel about Life and its Correlations with Other Variables in General Social Survey 2017

### 2022-03-22 outputs/paper/paper.Rmd
final final

The remainder of the paper is as follows: Section \@ref(data) performs the analyses using plots and tables and Section \@ref(results) explains the conclusion reached by the data. Following the conclusion there are three discussion points in Section \@ref(health,-mental-health-and-happiness), Section \@ref(economic-status-and-happiness), and Section \@ref(old-age-and-happiness). Then the paper goes over weaknesses and ideas for next steps in Section \@ref(weaknesses-and-next-steps).

@@

Out of the variables we compared to feelings about life, 'Self Rated Health' and 'Self Rated Mental Health' showed the most amount of visible correlation. People with better self-perceived health and mental health generally had more positive outlook in life, whereas poor conditions of both variables reported a negative outlook. Economical status also displayed a detectable correlation, with higher family income produced less negative feelings. Age and sex did not produce any perceptible relations, although higher age showed a minor increase in feelings on life.

@@

## Health, Mental Health and Happiness

DOes the older population have more positive outlook in life>

@@

Economic levels and happniess

@@

Health and happniess

@@

# Additional details

>>>>

The remainder of the paper is as follows: Section \@ref(data) performs the analyses using plots and tables and Section \@ref(results) explains the conclusion reached by the data. Following the conclusion there are three discussion points in Section \@ref(health-and-happiness), Section \@ref(economic-status-and-happiness), and Section \@ref(old-age-and-happiness). Then the paper goes over weaknesses and ideas for next steps in Section \@ref(weaknesses-and-next-steps).

@@

Out of the variables we compared to feelings about life, 'Self Rated Health' and 'Self Rated Mental Health' showed the most amount of visible correlation. People with better self-perceived health and mental health generally had more positive outlook in life, whereas poor conditions of both variables reported a negative outlook. Economical status also displayed a detectable correlation, with higher family income produced less negative feelings. Age and sex did not produce any significant relations, although higher age showed a minor increase in feelings on life.

@@

## Health and Happiness

To further analyze the relationship between physical and mental health and feelings about life, we will focus on one crucial aspect of positive feelings: happiness. Since we do not know the causal relation between the two variables, we need to be open to all of the options. One is that better health increases happiness. This is possible as negative emotions develop from feelings of powerlessness and pain. The other option is that happiness improves health. In many studies, happiness is seen to lower your blood pressure, improve sleep, lowers the risk of cardiovascular diseases, and reduce stress [@citehealth]. Optimism and positivity are also known to be linked to the overall well-being of a person, and these reasons may be why we saw that the health had clear relationship with overall feelings about life.

@@

Does money buy happiness? This is a question often debated from scholarly articles to dinner tables. In one study in 2010, it was reported that happiness and wealth had direct relationship and happiness increased with income, but only up to about 75,000 dollars a year [@gaetano]. After the salary of 75,000 dollars, happiness did not increase thus a person making 100,000 dollars a year had similar levels of happiness as someone making 75,000 dollars a year. This is thought to be because although wealth and happiness does not have a causal relation, wealth removes a lot of financial stress which in turn increases satisfaction. In another study, income produced higher evaluation of life, but had no relationship with emotional well-being, which aligns with our outcomes [@kahneman].

@@

Last discussion point asks why there was an uptick in feelings about life with old age, whereas no trend was found with younger age. One possible explanation of this is related to health as discussed above. If happier people tend to be more healthy, then they will live longer on average, thus increasing the mean score for old age. One study found that happier people had a greater chance of living longer, and had greater chance living past 85 years of age [@topor]. It is plausible that these two variables do not have a direct causal relation but have a common causal variable such as health.

@@

# Supplementary Survey

A supplementary survey has been made to further assess the relationship inquired in the paper [here](https://forms.gle/dVZN2XFPy5deFeXx6).

Direct URL: https://forms.gle/dVZN2XFPy5deFeXx6

This survey aims to find relationships of Canadian residents' feelings about life with various variable such as age, economical status, and self rated health. Feel free to answer 'N/A' on any questions you feel uncomfortable answering.

There are 10 questions:

  - Age
  - Sex
  - Self rated physical health
  - Self rated mental health
  - History of medication
  - History of illness
  - Household income
  - Disposable income
  - Disposable savings
  - Feelings about life as a whole.
